---
title: "Illustrative case study 2"
author: "Thomas Debray"
date: "This report was generated on `r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---


```{r setup, include=FALSE}
# Set document properties
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

source("../R/functions.r")

library(table1)
library(dplyr)
library(ggplot2)
library(rms)
```


```{r}
################################################################################
# Prepare data
################################################################################

# Fetch IST v2.0 from the web
d <- read.csv("https://datashare.ed.ac.uk/bitstream/handle/10283/124/IST%20dataset%20supp1%20%281%29%20trials%202011.csv?sequence=4&isAllowed=y")

dat <- data.frame(HOSPNUM = d$HOSPNUM,
                  OCCODE = d$OCCODE,
                  RXASP = d$RXASP,
                  RXHEP = d$RXHEP,
                  RDELAY = d$RDELAY,
                  AGE = d$AGE,
                  SEX = d$SEX,
                  RSLEEP = d$RSLEEP,
                  RCONSC = d$RCONSC,
                  RATRIAL = d$RATRIAL,
                  RSBP = d$RSBP,
                  STYPE = d$STYPE,
                  RDEF1 = d$RDEF1,
                  RDEF2 = d$RDEF2,
                  RDEF3 = d$RDEF3,
                  RDEF4 = d$RDEF4,
                  RDEF5 = d$RDEF5,
                  RDEF6 = d$RDEF6,
                  RDEF7 = d$RDEF7,
                  RDEF8 = d$RDEF8,
                  RVISINF = d$RVISINF,
                  RCT = d$RCT,
                  RASP3 = d$RASP3,
                  RHEP24 = d$RHEP24,
                  EXPDD = d$EXPDD)

# Exclude patients
dat_ASP <- dat %>% filter(RXASP == "Y", # Only include Aspirin patients
                          OCCODE %in% c(1, 2, 3, 4),  # Only include patients with complete 6mo outcome data
                          RASP3 %in% c("Y", "N"),
                          RATRIAL %in% c("Y","N"),
                          RHEP24 %in% c("Y", "N")) %>%
  mutate(Trt = ifelse(RDELAY <= 8, 1, 0),
         outcome = ifelse(OCCODE %in% c(1,2), 1, 0)) # Combine different outcome states
```

The International Stroke Trial (IST) was a large, multi-center, randomized, placebo-controlled trial including `r nrow(dat)` stroke patients. Individual participant data are available from [Edinburgh DataShare](https://datashare.ed.ac.uk/handle/10283/124). The IST evaluated the effect of Aspirin on a primary composite outcome of death or dependency (i.e. absence of autonomy) at 6 months (binary outcome). In this illustrative case study, we focus on `r nrow(dat_ASP)` patients from IST database version 2 that were randomized to receive Aspirin. 

We excluded patients with

* Missing data for the primary outcome (N = `r length(which(dat_ASP$OCCODE == 9 | dat_ASP$OCCODE == 0))`)
* No information on whether Aspirin was administered within 3 days prior to randomisation  (N = `r length(which(!(dat_ASP$RASP3 %in% c("Y", "N"))))`)
* No information on whether Heparin was administered within 24 hours prior to randomisation (N = `r length(which(!(dat_ASP$RHEP24 %in% c("Y", "N"))))`)
* No information on the presence of atrial fibrillation (N = `r length(which(!(dat_ASP$RATRIAL %in% c("Y", "N"))))`)

```{r}
dat_ASP$y <- dat_ASP$outcome

dat_ASP$outcome <- factor(dat_ASP$outcome, 
                          levels = c(1,0), 
                          labels = c("Death or dependency at 6 months", "(not) recovered"))
```

Amongst patients enrolled in the Aspirin arm of the International Stroke Trial, less than a quarter received the treatment between 0 and 8 hours following the onset of symptoms (`r round(nrow(subset(dat_ASP, Trt == 1))/nrow(dat_ASP)*100,1)`%, N = `r nrow(subset(dat_ASP, Trt == 1))`). In this group, `r nrow(subset(dat_ASP, Trt == 1 & y == 1))` (`r round(nrow(subset(dat_ASP, Trt == 1 & y == 1))/nrow(subset(dat_ASP, Trt == 1))*100,1)`%) patients were recorded with death or dependency at 6 months versus `r nrow(subset(dat_ASP, Trt == 0 & y == 1))` (`r round(nrow(subset(dat_ASP, Trt == 0 & y == 1))/nrow(subset(dat_ASP, Trt == 0))*100,1)`%) in the group of patients who received Aspirin later (after 8 hours).

We here consider the following baseline covariates (some of which were imputed):

```{r}
dat_ASP$albl <- factor(dat_ASP$Trt, 
                    levels = c(1,0), 
                    labels = c("Early Aspirin administration", "Late Aspirin administration"))
dat_ASP$SEX <- factor(dat_ASP$SEX, 
                    levels = c("F","M"), 
                    labels = c("Female", "Male"))
dat_ASP$RCONSC <- factor(dat_ASP$RCONSC, 
                    levels = c("F","D", "U"), 
                    labels = c("Fully alert", "Drowsy", "Unconscious"))
dat_ASP$RCT <- factor(dat_ASP$RCT, 
                    levels = c("N", "Y"), 
                    labels = c("No", "Yes"))
dat_ASP$RVISINF <- factor(dat_ASP$RVISINF, 
                    levels = c("N", "Y"), 
                    labels = c("No", "Yes"))
dat_ASP$STYPE <- factor(dat_ASP$STYPE, 
                    levels = c("PACS", "TACS", "POCS", "LACS", "OTH"), 
                    labels = c("PACS", "TACS", "POCS", "LACS", "Other"))

label(dat_ASP$AGE) <- "Age"
label(dat_ASP$RSBP) <- "Systolic blood pressure at randomisation"
label(dat_ASP$SEX) <- "Gender"
label(dat_ASP$RCONSC) <- "Conscious state at randomisation"
label(dat_ASP$RCT) <- "CT before randomisation"
label(dat_ASP$RVISINF) <- "Infarct visible on CT"
label(dat_ASP$STYPE) <- "Stroke subtype"


units(dat_ASP$AGE) <- "years"
units(dat_ASP$RSBP) <- "mmHg"


table1(~ AGE + RSBP + SEX + RCONSC + RCT + RVISINF
       + STYPE| albl, 
       data = dat_ASP, 
       caption = "Baseline characteristics of the patients that were administered Aspirin in the IST.")
```

# Estimation of the treatment effect
```{r}
# We first estimate the prognostic score
set <- 'rcs(AGE,3) + rcs(RSBP,3) + SEX + RCONSC + RCT + RVISINF + STYPE + 
RDEF1 + RDEF2 + RDEF3 + RDEF4 + RDEF5 + RDEF6 + RDEF7 + RDEF8 + RATRIAL + RASP3'

fmla.g <- formula(paste("y~", set))
pgs <- glm(fmla.g, data = dat_ASP[dat_ASP$Trt == 0,], family = "binomial")
dat_ASP$pgs <- predict(pgs, newdata = dat_ASP, "response")
```

We consider the Expanded Disability Status Scale (EDSS) measured 36 weeks after baseline as the primary outcome. Subsequently, we estimate the ATT in the simulated observation study using five approaches:

* Naive: A linear regression ignoring the potential presence of confounding and effect modification.
* DSR NNM: nearest neighbour matching using the disease risk score
* DSR OFM: optimal full matching using the disease risk score
* DSR IPW: inverse probability weighting using the disease risk score
* DSR TDW: target distribution weighting using the disease risk score

The disease risk score was calculated using linear regression in the control group, and includes the following baseline covariates as main effects: [TODO]

The distribution of the disease risk score is as follows:

```{r, fig.height = 5, fig.width = 10}
data <- dat_ASP %>%
    spread(Trt, "pgs", sep = "_p")
  
  ggplot(data) + 
    geom_histogram(aes(x = Trt_p1, fill = "Early Aspirin administration"), color = "red", alpha = 0.5) + 
    geom_histogram(aes(x = Trt_p0, y = -..count.., fill = "Late Aspirin administration"), color = "blue", alpha = 0.5) + 
    ylab("Number of patients") + 
    xlab("Predicted risk of death or dependence at 6 months") +
    xlim(0,1) + 
    geom_hline(yintercept = 0, lwd = 0.5) +
    scale_y_continuous(labels = abs) +
    theme(legend.position = "bottom", legend.title = element_blank()) +
    ggtitle("Distribution of the disease risk score") +
    scale_color_brewer(palette = "Set1") + 
    scale_fill_brewer(palette = "Set1")
```


# Results

```{r}
results <- analyze_nrs(dat_ASP)

results <- data.frame("Method" = results$method,
                      "Estimate" = format(round(results$est_beta, 2), nsmall = 2),
                      "SE" = format(round(results$est_se, 2), nsmall = 2),
                      "95%CI" = paste(format(round(results$est_beta_CIl, 2), nsmall = 2), "; ", format(round(results$est_beta_CIu, 2), nsmall = 2)),
                      "Time" = format(round(results$est_time, 2), nsmall = 2))

knitr::kable(results)
```

Finally, we apply bootstrapping to improve estimation of standard errors:

```{r}
results <- analyze_nrs_bs(dat_ASP, fn = "bootstrap_cs2.rda")
results <- results %>% group_by(method) %>% summarise(beta = mean(est_beta), 
                                                      beta_se = mean(est_se),
                                                      beta_CIl = quantile(est_beta, 0.025),
                                                      beta_CIu = quantile(est_beta, 0.975),
                                                      time = sum(est_time))

results <- data.frame("Method" = results$method,
                      "Reference" = 0,
                      "Estimate" = format(round(results$beta, 2), nsmall = 2),
                      "SE" = format(round(results$beta_se, 2), nsmall = 2),
                      "95%CI" = paste(format(round(results$beta_CIl, 2), nsmall = 2), "; ", format(round(results$beta_CIu, 2), nsmall = 2)),
                      "Time" = format(round(results$time, 2), nsmall = 2))
knitr::kable(results)
```



