---
title: "Report"
author: "Thomas Debray"
date: "7-12-2021"
output: html_document
---

```{r setup, include=FALSE}
# Set document properties
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

source("../R/functions.r")

library(ggplot2)

################################################################################
# Prepare data
################################################################################
dat <- load_imputed_data("F:/Projects/Datasets/Weights disease risk score/IDB_control_imputed.rda")

dat_baseline <- dat %>% 
  group_by(USUBJID.x, TRIAL) %>% 
  filter(AVISITN == 0) %>%
  summarize(
    AGE = first(AGE),
    SEX = first(MaleGender),
    RACE = first(WhiteRace),
    WEIGHTBL = first(WEIGHTBL),
    ONSYRS = first(ONSYRS),
    DIAGYRS = first(DIAGYRS),
    RLPS1YR = first(RLPS1YR),
    RLPS3YR = first(RLPS3YR),
    EDSSBL = first(EDSSBL),
    GDLESBL = first(GDLESBL),
    T1LESBL = first(T1LESBL),
    T2LESBL = first(T2LESBL),
    PRMSGR = first(PRMSGR),
    NHPTMBL = first(NHPTMBL),
    PASATABL = first(PASATABL),
    T25FWABL = first(T25FWABL),
    TRELMOS = first(TRELMOS)
)

npat <- nrow(dat_baseline)
```

# Original data sources
Previously, four RCTs of relapsing-remitting multiple sclerosis (RRMS) were combined in an integrated clinical trial database. This database includes patient-level data from ADVANCE (pegylated interferon beta-1a s.c. registration trial), DEFINE and CONFIRM (dimethyl fumarate registration trials), and AFFIRM (natalizumab registration trial). Details of these trials are provided below:

* The ADVANCE trial includes `r  length(unique(subset(dat, TRIAL == "ADVANCE" & TRTA == "Placebo")$USUBJID.x))` patients who received placebo every 2 weeks for 48 weeks followed by 125 $\mu$g peginterferon beta-1a subcutaneously every 2 or 4 weeks for 48 weeks (https://www.clinicaltrials.gov/ct2/show/NCT00906399) 
* The DEFINE trial includes `r length(unique(subset(dat, TRIAL == "DEFINE" & TRTA == "Placebo")$USUBJID.x))` patients who received two placebo capsules orally three times daily (https://www.clinicaltrials.gov/ct2/show/NCT00420212)
* The CONFIRM trial includes `r length(unique(subset(dat, TRIAL == "CONFIRM" & TRTA == "Placebo")$USUBJID.x))` patients who received two placebo capsules orally three times daily (https://www.clinicaltrials.gov/ct2/show/NCT00451451) 
* The AFFIRM trial includes  `r length(unique(subset(dat, TRIAL == "AFFIRM" & TRTA == "Placebo")$USUBJID.x))` patients who received placebo, intravenous infusion, every 4 weeks, for up to 116 weeks.  (https://www.clinicaltrials.gov/ct2/show/NCT00027300)

We here consider the  following baseline covariates:

* AGE - age (years)
* SEX - gender (1 = male, 0 = female)
* RACE - ethnicity, transformed as (1 = white, 0 = non-white)
* WEIGHTBL - weight (kg)
* ONSYRS - Time since first multiple sclerosis symptoms (years)
* DIAGYRS - Time since multiple sclerosis diagnosis (years)
* EDSSBL - Baseline EDSS  (discrete)
* PRMSGR - Prior MS Treatment Group (1 = yes, 0 = no)
* NHPTMBL - Baseline 9 Hole Peg Test Average Score
* PASATABL -	Baseline PASAT 3 
* T25FWABL - Baseline Timed 25 Foot Walk 
* Relapses
  * RLPS1YR - No. of Relapses within the previous 12 months 
  * RLPS3YR - No. of Relapses within the previous 3 years 
  * TRELMOS - No. of months Since Recent Pre-Study Relapse
* MRI findings
  * GDLESBL - Number of Gd+ lesions 
  * T1LESBL - Number of T1 weighted lesions 
  * T2LESBL - Number of T2 weighted lesions 


```{r, fig.height = 20, fig.width = 10}
ggdat <- data.frame(x = c(dat_baseline$AGE, 
                          dat_baseline$WEIGHTBL, 
                          dat_baseline$ONSYRS,
                          dat_baseline$DIAGYRS,
                          dat_baseline$EDSSBL,
                          dat_baseline$RLPS1YR,
                          dat_baseline$RLPS3YR,
                          dat_baseline$GDLESBL,
                          dat_baseline$T1LESBL,
                          dat_baseline$T2LESBL,
                          dat_baseline$NHPTMBL,
                          dat_baseline$PASATABL,
                          dat_baseline$T25FWABL,
                          dat_baseline$TRELMOS
                          ),
                    var = c(rep("Age (years)", npat), 
                            rep("Weight (kg)", npat), 
                            rep("Time since first symptoms (years)", npat),
                            rep("Time since diagnosis (years)", npat),
                            rep("Baseline EDSS", npat),
                            rep("No. of Relapses (12mo)", npat),
                            rep("No. of Relapses (36mo)", npat),
                            rep("No. of Gd+ lesions", npat),
                            rep("No. of T1 lesions", npat),
                            rep("No. of T2 lesions",npat),
                            rep("9 Hole Peg Test Average Score", npat),
                            rep("PASAT 3", npat), 
                            rep("Timed 25 Foot Walk", npat),
                            rep("No. of months Since Recent Pre-Study Relapse",npat)
                            ),
                    Study = rep(dat_baseline$TRIAL, 14))

ggplot(ggdat, aes(x = x, color = Study)) + geom_density() + 
  facet_wrap( ~ var, scales = "free", ncol = 2) +
  xlab("") +
  theme(legend.position = "bottom")
```


# Hypothetical observational study
We generated a hypothetical observational study by combining the placebo arms from aforementioned four clinical trials. The primary outcome for each study is the Expanded Disability Status Scale (EDSS) measured 36 weeks after baseline. The EDSS score is a scale that quantifies disability in 0.5 unit increments, and ranges from 0 (no disability) to 10 (death due to MS).

```{r}
# Prepare NRS dataset
dat_NRS <- prepare_nrs(dat, 
                       STUDYID_control = c("ADVANCE"), 
                       STUDYID_treat = c("AFFIRM", "DEFINE", "CONFIRM"), 
                       SUBGROUP = TRUE)
```


